### 2024-10-10 paper/paper.qmd
Update

---
title: "Spotify Data"
author: "Yanzun Jiang, "
date: today
date-format: long
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# load packages and read data
library(tidyverse)
library(knitr)
```

\newpage

# Introduction

To conduct the analysis of songs on Spotify, a dataset downloaded from @spotify was utilized, as described in @sec-data. Based on the initial findings, it was observed that `tempo` and `valence` are not positively correlated in a simple (linear) way, while energy falls then rises with time (@sec-discussion). Also, supplementary insights are provided in @sec-appendix.

# Data {#sec-data}

```{r}
#| include: false
#| warning: false
#| message: false

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
```

## Overview

The dataset used in this analysis is the songs of David Tao sourced from Spotify (@spotify). It records various aspects of the songs by David from 1997 to present.

The variables analyzed in this study are listed below:

`album_release_date`: the date that the album was released.

`energy`: the energy of the song.

`valence`: the degree of happiness of the song.

`tempo`: the pace of the song.

The dataset was accessed using the `spotifyr` package (@spotifyr). For the analysis, the R programming language was employed (@r), utilizing the `tidyverse` (@tidyverse) package for data cleaning, transformation, visualization, the `knitr` package to visualize the data, and the `here` package (@here) accessing the data in this analysis. Afterward, the cleaned dataset was processed and tested using additional functions from the `tidyverse` package (@tidyverse).

## Preview of the Dataset

This (@tbl-data) is the preview of the dataset.

```{r}
#| label: tbl-data
#| tbl-cap: Preview of Data
#| echo: false
#| warning: false
#| message: false

kable(head(data), format = "markdown")
```

# Discussion {#sec-discussion}

```{r, fig.width=4, fig.height=2}
#| label: fig-tem-val
#| fig-cap: Relationship of Tempo and Valence in David's Songs
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = tempo, y = valence)) +
  geom_point(col = "lightblue") +
  geom_smooth() +
  labs(x = "Tempo", y = "Valence")
```

```{r, fig.width=4, fig.height=2}
#| label: fig-time-ener
#| fig-cap: Changes of Energy with Time in David's Songs
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = album_release_date, y = energy)) +
  geom_point(col = "lightblue") +
  geom_smooth() +
  labs(x = "Date Released", y = "Energy")
```

From @fig-tem-val, we can see that `tempo` and `valence` are not strictly positively correlated as we expected. This might due to the lyric and usage of different musical instruments.

From @fig-time-ener, we can find that the energy for David's Songs falls then rises with time, which might be associated with his personal experience.

\newpage

\appendix

# Appendix {#sec-appendix}

## Data Cleaning

The data cleaning process involved tidying the dates, filtering out useless columns from the raw dataset, and filtering out observations with NAs.

## Summary Statistics of the Data

```{r}
#| label: tbl-energy
#| tbl-cap: Summarise Statistics of Energy in David's Songs
#| echo: false
#| warning: false
#| message: false

summarise_energy <- data |>
  summarise(Mean = round(mean(energy), 3),
            Median = round(median(energy), 3),
            Variance = round(var(energy), 3),
            Min = round(min(energy), 3),
            Max = round(max(energy), 3),
            IQR = round(IQR(energy), 3)
  )
kable(summarise_energy, format = "markdown")
```

```{r}
#| label: tbl-valence
#| tbl-cap: Summarise Statistics of Valence in David's Songs
#| echo: false
#| warning: false
#| message: false

summarise_valence <- data |>
  summarise(Mean = round(mean(valence), 3),
            Median = round(median(valence), 3),
            Variance = round(var(valence), 3),
            Min = round(min(valence), 3),
            Max = round(max(valence), 3),
            IQR = round(IQR(valence), 3)
  )
kable(summarise_valence, format = "markdown")
```

```{r}
#| label: tbl-tempo
#| tbl-cap: Summarise Statistics of Tempo in David's Songs
#| echo: false
#| warning: false
#| message: false

summarise_tempo <- data |>
  summarise(Mean = round(mean(tempo), 3),
            Median = round(median(tempo), 3),
            Variance = round(var(tempo), 3),
            Min = round(min(tempo), 3),
            Max = round(max(tempo), 3),
            IQR = round(IQR(tempo), 3)
  )
kable(summarise_tempo, format = "markdown")
```

The summary statistics provides key statistical insights into the variables of David's songs. For each variable, the summarize tables (@tbl-energy, @tbl-valence, and @tbl-tempo) reports the mean, median, variance, minimum value, maximum value, and interquartile range of those variables, offering a concise overview of range, central tendencies, and variability in the data.

\newpage

# References